---
title: "GSS challange More Data"
author: "David Barton"
date: " October 2025"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

This problem is a categorical problem. After looking into it, we found a problem where there are three different factors. White, black, and in between. Everyone else is in a category called other. Cappun has two different variables: those who are for and those who are against.   

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(cappun,race) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=cappun,y=count,fill=race))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

What the graph shows is that there are more people in the white range than in any other. What is most telling is that both branches are similar when it comes to white people and their support for it. It is easy to see that the black branch is not in favor of capital punishment. When you put all the branches they all add to 100. 

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~cappun + race, data=mydata)
rowPerc(table2)
colPerc(table2)
```

After looking at the number, out of all the people who favor capital punishment, eighty percent of them are white. while nine percent where black. Still, a majority of whites opposed capital punishment. That being sixty-five percent. while 42 percent of blacks opposed it. 

## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->




## More data version

```{r}
fisherXY(mydata$race,mydata$cappun)
```

The P-value from comparing both is very low, showing there is something different between white people and other people when it comes to supporting and opposing capital punishment. Even with a 95% confidence level. Each group, compared to the others, is different. Different people disagree with each other. Each group is different.  

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

This branch is very similar to the second branch. The only difference here is that there is a statistical difference between the whites and the blacks, and everyone else. On who is for and is against capital punishment. Everyone else is the same. While in the second branch, the statistical difference was between everyone, the whites, and the blacks. The whites were for capital punishment, and they were the majority group who were also against it. While the majority of blacks were not in favor of the capital punishment. There were some who were in favor but it was highly unlikely that they were. After I saw the P-value was really low for everyone it only proved the case that they are different. That was not the case for the second branch as only the whites and the blacks had a low P-value and the other ones didn't. 
